---
title: Lab Eight -- Probability # Title of the document

# Bibliography Details
bibliography: Lab8References.bib  # Name of bib file in the same folder 
bibliographystyle: apa            # Bibliography style

execute:               # Default settings for code blocks
  warning: false       # Do not print warnings
  message: false       # Do not print messages
  size: "scriptsize"   # Set code size to script size

# Below will do html by default. To render the entire project
# Use the following code:
# quarto::quarto_render("Lab8WriteUp.qmd")
format:
  pdf:        # Specify .pdf formatting options 
    documentclass: article     # Document Class
    number-sections: true      # Number Sections
    reference-section-title: "References" # Title for reference section
    code-line-numbers: true    # Line numbers printed in code
    cap-location: top          # Captions at the top of tables 
    geometry:
      - top=0.75in
      - left=0.75in
      - bottom=0.75in
      - right=0.75in
  html:       # Specify HTML formatting options 
    number-sections: true      # Number Sections
    code-fold: true            # Button to show code
    code-line-numbers: true    # Line numbers printed in code
    cap-location: top          # Captions at the top of tables and figures
format-links: [pdf]
---
```{r}
#| echo: false

default_chunk_hook  <- knitr::knit_hooks$get("chunk")

latex_font_size <- c("Huge", "huge", "LARGE", "Large", 
                     "large", "normalsize", "small", 
                     "footnotesize", "scriptsize", "tiny")

knitr::knit_hooks$set(chunk = function(x, options) {
  x <- default_chunk_hook(x, options)
  if(options$size %in% latex_font_size) {
    paste0("\n \\", options$size, "\n\n", 
      x, 
      "\n\n \\normalsize"
    )
  } else {
    x
  }
})
```

```{r}
#| label: load-packages
#| include: false
library(tidyverse)
library(patchwork)
```

**Tips**

* Complete the tasks below. Make sure to start your solutions in on a new line that starts with "**Solution**:".
* Make sure to use the Quarto Cheatsheet. This will make completing and writing up the lab *much* easier.
* Make sure to use "enter" to make your code readable, so it doesn't run off the page. I switched the Quarto document to automatically render to .pdf, so you can see what you're submitting by default. You can switch back for the highlighting by moving the html block above the pdf block in the YAML header.
* Don't forget that you can copy-and-paste code when you're doing something similar as we did in class!
* Make sure to plot all computed probabilities. This is good `ggplot` practice AND will help make sure you don't make a mistake when computing the probability.

# The Binomial and Negative Binomial Distributions

## The Binomial Distribution
In many introductory courses, we make the simplifying assumption of independence in Bernoulli trials. For example, when playing best two-out-of-three legs in a darts match, we would assume Player A has an 70% chance of winning each leg. We write:
$$X \sim \text{Binomial}(n=3, p=0.70)$$
where

* $X$ is the number of legs won
* $n$ is the number of trials 
* $p$ is the probability of the event of interest

Of note, it is not necessarily the case that the players play all three legs -- if either player wins legs one and two the match ends.

There are eight possible outcomes for Player A:

* (win, win, "win") $\longrightarrow$ Player A wins
* (win, win, "lose") $\longrightarrow$ Player A wins
* (win, lose, win) $\longrightarrow$ Player A wins
* (win, lose, lose) $\longrightarrow$ Player A loses
* (lose, win, win)  $\longrightarrow$ Player A wins
* (lose, win, lose) $\longrightarrow$ Player A loses
* (lose, lose, "win") $\longrightarrow$ Player A loses
* (lose, lose, "lose") $\longrightarrow$ Player A loses

**Note**: I use quotations above to denote outcomes where the last match isn't actually played.

**Use the Binomial distribution to compute the probability that Player A wins the match.  Make sure write out your probability statement and work to write it in terms of the distribution function (PMF/CDF) to plot the probability using `ggplot`.**

**Solution**
```{r}
#| eval: false

install.packages("tidyverse")
```
```{r}
library("tidyverse")
```

The probability that Player A wins the match is the probability of winning two or more legs.

$P(X\geq2) = f_X(2) + f_X(3) = 0.784$

```{r}
(dbinom(x=2, prob = 0.70, size = 3) + dbinom(x=3, prob = 0.70, size = 3))
```

```{r}
#| label: fig-bnom
#| fig-cap: "The probability that Player A wins the match"
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x = 0:3) |>
  mutate(PMF = dbinom(x, prob = 0.70, size = 3),
         CDF = pbinom(x, prob = 0.70, size = 3))

PMF.plot <- ggplot() +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_linerange(data = ggdat |> filter(x >= 2),
                 aes(x = x, ymin = 0, ymax = PMF),
                 color = "darkorchid2", linewidth = 2, alpha = 0.50) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylim(0, 1) +
  ylab("Probability") +
  theme_bw()

CDF.plot <- ggplot() +
  geom_step(data = ggdat, aes(x = x, y = CDF)) +
  geom_point(data = ggdat, aes(x = x, y = CDF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylab("Cumulative Probability") +
  theme_bw()

PMF.plot | CDF.plot
```

## The Negative Binomial Distribution
In intermediate courses, we learn the Negative Binomial distribution that we use to model the number of "failures" until the $r$th success. Note that this approach still makes the simplifying assumption of independence in Bernoulli trials.

**Specify the two cases necessary to compute the probability. Fully explain these probability statements in the context of a best two-out-of-three match and the Negative Binomial distribution. Finally, compute the probability that Player A wins the match and show you get the same result as the Binomial approach. Make sure write out your probability statement and work to write it in terms of the distribution function (PMF/CDF) to plot the probability using `ggplot`.**

**Solution**

The two cases where Player A wins is when they have 1 loss or no loss before their 2nd success: (lose, win, win) or (win, lose, win).

$P(X<2) = f_X(1) + f_X(0) = F_X(1) = 0.784$

```{r}
(dnbinom(x=1, prob = 0.70, size = 2)+ dnbinom(x=0, prob = 0.70, size = 2))
```

```{r}
#| label: fig-nbnom
#| fig-cap: "The probability that Player A wins the match"
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x = 0:3) |>
  mutate(PMF = dnbinom(x, prob = 0.70, size = 2),
         CDF = pnbinom(x, prob = 0.70, size = 2))

PMF.plot <- ggplot() +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_linerange(data = ggdat |> filter(x < 2),
                 aes(x = x, ymin = 0, ymax = PMF),
                 color = "darkorchid2", linewidth = 2, alpha = 0.50) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:2) +
  ylab("Probability") +
  theme_bw()

CDF.plot <- ggplot() +
  geom_step(data = ggdat, aes(x = x, y = CDF)) +
  geom_point(data = ggdat, aes(x = x, y = CDF)) +
  geom_point(data = ggdat |> filter(x==1),
            aes(x = x, y = CDF),
            color="darkorchid2", size=2.5, alpha=0.50) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylab("Cumulative Probability") +
  theme_bw()

PMF.plot | CDF.plot
```

# The Beta Binomial Distribution
One approach to generalize this modeling is to use a Beta-Binomial distribution. This distribution arrises when we let $p$ be a random variable that follows a Beta distribution:
\begin{align*}
  P &\sim \text{Beta}(\alpha, \beta)\\
  X &\sim \text{Binomial}(n, P).
\end{align*}
This distribution is not part of base `R`, but is available using the `_bbinom` functions from the `extraDistr` package [@extraDistr].

The primary benefit is that the Beta-Binomial enables us to specify some uncertainty about the skill of Player A, as $p$ does not have to be fixed (e.g., $p=0.70$). Suppose the result of the best two-out-of-three match follows a Beta-Binomial distribution  where $\alpha=7$ and $\beta=3$.

## The Beta Part
1. **Plot the corresponding Beta distribution. Interpret the plot in the context of a best two-out-of-three match and the Binomial distribution.**\
**Hint**: The Beta distribution is available in base `R` via `_beta()`.

**Solution**
```{r}
#| eval: false
install.packages("extraDistr")
```

```{r}
library("extraDistr")
```

```{r}
# We're specifying a function for the probability of winning 1 leg rather than a number. 
# x is the probability of winning 1 leg. We're random-sampling so there's no size in dbeta and pbeta.

#| label: fig-beta
#| fig-cap: "The probability that Player A wins the match"
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x = seq(0, 1, length.out=1000))  |>
  mutate(PDF = dbeta(x, 7, 3, 
                      log = FALSE),
        CDF = pbeta(x, 7, 3, lower.tail = TRUE, 
                      log.p = FALSE))

PDF.plot <- ggplot() +
  geom_line(data = ggdat,
                 aes(x = x, y = PDF)) +
  geom_hline(yintercept = 0) +
  ylab("Probability") +
  theme_bw()

CDF.plot <- ggplot() +
  geom_line(data = ggdat, aes(x = x, y = CDF)) +
  geom_hline(yintercept = 0) +
  ylab("Cumulative Probability") +
  theme_bw()

PDF.plot | CDF.plot
```

The probability of Player A winning a leg cluster around 0.70, meaning on average, Player A still wins 70% of the time. The distribution is left-skewed, i.e.,there are more higher values than lower values, meaning the probability of Player A winning 1 leg is generally higher around $0.70-0.80$.

2. How does the Beta incorporate uncertainty about the skill of Player A? **Compute the mean and variance of the distribution.**\
**Hint**: You will find the `integrate(...)` function helpful here. 
\begin{align*}
  E(P) &= \mu_P = \int_{0}^1 p f_P(p) dp\\
  var(P) &= \sigma^2_P = \int_{0}^1 (p-\mu_P)^2 f_P(p) dp\\
\end{align*}

**Solution**
The Beta incorporate uncertainty about the skill of Player A because it treats the probability of winning as a variable with a distribution (rather than a constant).

$E(P)=0.7, var(P)=0.019$
```{r}
integrand_mean = function(p) {
  p*(dbeta(p, 7, 3, log = FALSE))
  }
(mean=integrate(integrand_mean, 0, 1))

integrand_variance = function(p) {
  ((p-mean$value)^2)*(dbeta(p, 7, 3, log = FALSE))
  }
(variance=integrate(integrand_variance, 0, 1))
```

3. **What is the probability that Player A has a win probability of exactly 0.70?  Make sure write out your probability statement and work to write it in terms of the distribution function (PDF/CDF) to plot the probability using `ggplot`.**\
**Hint:** Think carefully about the properties of continuous distributions here.

**Solution**

The probability of Player A having a win probability of exactly 0.70 is 0. Under a continuous distribution, a unit can always be made finer (i.e., a variable can take on an infinite number of possible values within any interval).

4. **What is the probability that Player A has a win probability of at least 0.50?  Make sure write out your probability statement and work to write it in terms of the distribution function (PDF/CDF) to plot the probability using `ggplot`.**

**Solution**

The probability of Player A having a win probability of at least 0.5 is 
$$P(p\geq0.5)=P(p>0.5)=1-F_x(0.5)=0.91$$

```{r}
(1-pbeta(0.5, 7, 3))
```

```{r}
#| label: fig-beta1
#| fig-cap: "The probability that Player A has a win probability of at least 0.5"
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x = seq(0, 1, length.out=1000))|>
  mutate(PDF = dbeta(x, 7, 3, 
                      log = FALSE),
        CDF = pbeta(x, 7, 3, lower.tail = TRUE, 
                      log.p = FALSE))

PDF.plot <- ggplot() +
  geom_line(data = ggdat,
                 aes(x = x, y = PDF)) +
  geom_ribbon(data = ggdat|> filter(x >=0.5),
              aes(x = x, ymin=0, ymax=PDF),
              fill="darkorchid2", alpha=0.50) +
  geom_hline(yintercept = 0) +
  ylab("Probability") +
  theme_bw()

CDF.plot <- ggplot() +
  geom_line(data = ggdat, aes(x = x, y = CDF)) +
  geom_hline(yintercept = 0) +
  ylab("Cumulative Probability") +
  theme_bw()

PDF.plot | CDF.plot

```

5. **What is the 90th percentile of Player A's win probability based on this Beta distribution? Make sure write out your probability statement and work to write it in terms of the distribution function (PDF/CDF) to plot the percentile using `ggplot`.**

**Solution**

The 90th percentile of Player A's win probability is $F_X^{-1}(0.9) = 0.87$.
```{r}
(qbeta(0.9, 7, 3))
```

```{r}
#| label: fig-beta2
#| fig-cap: "The 90th percentile of Player A’s win probability"
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x = seq(0, 1, length.out=1000))|>
  mutate(PDF = dbeta(x, 7, 3, 
                      log = FALSE),
        CDF = pbeta(x, 7, 3, lower.tail = TRUE, 
                      log.p = FALSE))

PDF.plot <- ggplot() +
  geom_line(data = ggdat, aes(x = x, y = PDF)) +
  geom_ribbon(data = ggdat |> filter(x <= qbeta(0.9, 7, 3)),
              aes(x = x, ymin = 0, ymax = PDF),
              fill = "darkorchid2", alpha = 0.5) +
  geom_hline(yintercept = 0) +
  ylab("Probability") +
  theme_bw()

CDF.plot <- ggplot() +
  geom_line(data = ggdat, aes(x = x, y = CDF)) +
  geom_point(data = tibble(x = qbeta(0.9, 7, 3))|> mutate(CDF = pbeta(x, 7, 3)),
            aes(x = x, y=CDF), color="darkorchid2")+
  geom_hline(yintercept = 0) +
  ylab("Cumulative Probability") +
  theme_bw()

PDF.plot | CDF.plot
```

## The Beta-Binomial Part
**Compute the probability that Player A wins the match. Compare your result to the Binomial and Negative Binomial results. Make sure write out your probability statement and work to write it in terms of the distribution function (PMF/CDF) to plot the probability using `ggplot`**

**Solution**

The probability of Player A winning the match is $P(X\geq2)=0.76$. This is lower than the results from the binomial and negative binomial distribution, because of the uncertainty about the skill of Player A.

```{r}
(dbbinom(2,3,7,3) + dbbinom(3,3,7,3))
```

```{r}
#| label: fig-bbnom
#| fig-cap: "The probability that Player A wins the match, considering uncertainty in skills"
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x = 0:3) |>
  mutate(PMF = dbbinom(x,3,7,3),
        CDF=pbbinom(x,3,7,3))

PMF.plot <- ggplot() +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_linerange(data = ggdat |> filter(x >= 2),
                 aes(x = x, ymin = 0, ymax = PMF),
                 color = "darkorchid2", linewidth = 2, alpha = 0.50) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylab("Probability") +
  theme_bw()

CDF.plot <- ggplot() +
  geom_step(data = ggdat, aes(x = x, y = CDF)) +
  geom_point(data = ggdat, aes(x = x, y = CDF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylab("Cumulative Probability") +
  theme_bw()

PMF.plot | CDF.plot
```


## The Why
Check the expected value and variance of each distribution.

Analytically, we have
\begin{align*}
  E(X) &= np  &&var(X) = np(1-p) \tag{Binomial Distribution}\\
  E(X) &= n\left(\frac{\alpha}{\alpha+\beta}\right)  &&var(X) = \frac{n \alpha \beta (\alpha + \beta + n)}{(\alpha+\beta)^2(\alpha + \beta + 1)} \tag{Beta-Binomial Distribution}
\end{align*}

**For each distribution, compute the mean number of legs won for player A.**\
**Hint**: You can perform the sum over the finite support set here. 
\begin{align*}
  E(X)   = \mu_X    &= \sum_{i=0}^3 i f_X(i)\\
  var(X) = \sigma_X &= \sum_{i=0}^3 (i-\mu_X) f_X(i)
\end{align*}

**Solution**

The two distributions have the same mean, but the Beta-Binomial Distribution has higher variance, which makes sense because it accounts for uncertainty in the skill of Player A.

* Binomial distribution: $E(X)=3\times0.70=2.1, var(X)=3\times0.70(1-0.70)=0.63$
* Beta-Binomial Distribution: $E(X)= 0.70(\frac{7}{7+3})=2.1, var(X)=\frac{3\times3\times(3+7+3)}{(3+7)^2(3+7+1)}=0.74$

Compute the mean of number of legs won for Player A.

* Binomial distribution:
```{r}
# sum
x=0:3
(mean = sum(x*dbinom(x, prob = 0.70, size = 3)))
(var=sum((x-mean)^2*dbinom(x, prob = 0.70, size = 3)))
```

* Beta-binomial distribution
```{r}
x=0:3
(mean = sum(x*dbbinom(x,3,7,3)))
(var=sum((x-mean)^2*dbbinom(x,3,7,3)))
```

# Altham's Multiplicative Binomial Distribution
While the Beta-Binomial enables us to model cases where winning has a clustering effect (e.g., a "hot hand").

## Probability Mass Function
The mass function for this distribution is below. While it used to be available in the `MBD` package for `R`, this package is no longer available.
$$f_X(x) = \frac{1}{c} {n \choose x} p^x (1-p)^{n-x} \theta^{x(n-x)}I(x \in \{0, 1, ..., n\})$$
where
$$c = \sum_{i=0}^n {n \choose i} p^i (1-p)^{n-i} \theta^{i(n-i)}$$ 

**Write a function in `R` called `daltbinom()` for computing the PMF of this distribution. Use your function to plot the distribution where $p=0.70$ where $\theta=0.50$, $\theta=1$, and $\theta=1.50$. Comment on the differences in the context of a best two-out-of-three match**

**Solution**

```{r}
daltbinom = function(x, n, p, theta) {
  i = 0:n
  c = sum(choose(n, i)*p^i*(1-p)^(n-i)*theta^(i*(n-i)))
  (1/c)*choose(n, x)*p^x*(1-p)^(n-x)*theta^(x*(n-x))
}
```

```{r}
#| label: fig-altbnom-pmf
#| fig-cap: "PMF of Altham's Multiplicative Binomial Distribution when $\\theta$=0.5, $\\theta$=1 and $\\theta$=1.5"
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x = 0:3) |>
  mutate(PMF = daltbinom(x,3,0.70,0.5))

PMF.plot.1 <- ggplot() +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylim(0,1)+
  ylab("Probability") +
  theme_bw()

ggdat <- tibble(x = 0:3) |>
  mutate(PMF = daltbinom(x,3,0.70,1))

PMF.plot.2 <- ggplot() +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylim(0,1)+
  ylab("Probability") +
  theme_bw()

ggdat <- tibble(x = 0:3) |>
  mutate(PMF = daltbinom(x,3,0.70,1.5))

PMF.plot.3 <- ggplot() +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylim(0,1)+
  ylab("Probability") +
  theme_bw()

PMF.plot.1 | PMF.plot.2 | PMF.plot.3

```

When $\theta=1$, it's a binomial distribution, which is a probability distribution of the number of leg Player A can win out of 3 trials that are independent. When $\theta$ is smaller than 1, the probability of Player A winning more legs after winning 1 leg is higher, so the probability of winning (WW"L" and WW"W") is significantly higher. When $\theta$ is larger than 1, Player A is more likely to win 2 trials out of 3 (2-1 wins are more likely than sweeps).

## Cumulative Distribution Function
**Use your `daltbinom()` function to create a `paltbinom()` function for computing the CDF of this distribution. Use your function to plot the distribution where $p=0.70$ where $\theta=0.50$, $\theta=1$, and $\theta=1.50$.**

**Solution**

The CDF at point x is just the sum of all the PMF values from 0 through x.
```{r}
# vectorized
paltbinom = function(x, n, p, theta) {
  sum(daltbinom(0:x, n, p, theta))
}
```

```{r}
#| label: fig-altbnom-cdf
#| fig-cap: "CDF of Altham's Multiplicative Binomial Distribution when $\\theta$=0.5, $\\theta$=1 and $\\theta$=1.5"
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x = 0:3) |>
  rowwise() |>
  mutate(CDF = paltbinom(x,3,0.70,0.5))

CDF.plot.1 <- ggplot() +
  geom_step(data = ggdat, aes(x = x, y = CDF)) +
  geom_point(data = ggdat, aes(x = x, y = CDF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylab("Cumulative Probability") +
  theme_bw()

ggdat <- tibble(x = 0:3) |>
  rowwise() |>
  mutate(CDF = paltbinom(x,3,0.70,1))

CDF.plot.2 <- ggplot() +
  geom_step(data = ggdat, aes(x = x, y = CDF)) +
  geom_point(data = ggdat, aes(x = x, y = CDF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylab("Cumulative Probability") +
  theme_bw()

ggdat <- tibble(x = 0:3) |>
  rowwise() |>
  mutate(CDF = paltbinom(x,3,0.70,1.5))

CDF.plot.3 <- ggplot() +
  geom_step(data = ggdat, aes(x = x, y = CDF)) +
  geom_point(data = ggdat, aes(x = x, y = CDF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylab("Cumulative Probability") +
  theme_bw()

CDF.plot.1 | CDF.plot.2 | CDF.plot.3

```

## The Winning Probability
**Compute the probability that Player A wins the match using Altham's Multiplicative Binomial Distribution. Compare your result to the Binomial and Negative Binomial results. Compare your results to the Beta-Binomial model, too. Make sure write out your probability statement and work to write it in terms of the distribution function (PMF/CDF) to plot the probability using `ggplot`.**

**Solution**

The probability that Player A wins the match is the probability of winning two or more legs. 

$$P(X\geq2) = 1-F_X(1)$$

For each distribution with different $\theta$, this probability is:
```{r}
(1-paltbinom (1,3,0.70,0.5))
(1-paltbinom (1,3,0.70,1))
(1-paltbinom (1,3,0.70,1.5))
```

```{r}
#| label: fig-altbnom-pmf1
#| fig-cap: "The probability that Player A wins the match when $\\theta$=0.5, $\\theta$=1 and $\\theta$=1.5"
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x = 0:3) |>
  mutate(PMF = daltbinom(x,3,0.70,0.5))

PMF.plot.1 <- ggplot() +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_linerange(data = ggdat |> filter(x >= 2),
                 aes(x = x, ymin = 0, ymax = PMF),
                 color = "darkorchid2", linewidth = 2, alpha = 0.50) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylim(0,1)+
  ylab("Probability") +
  theme_bw()

ggdat <- tibble(x = 0:3) |>
  mutate(PMF = daltbinom(x,3,0.70,1))

PMF.plot.2 <- ggplot() +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_linerange(data = ggdat |> filter(x >= 2),
                 aes(x = x, ymin = 0, ymax = PMF),
                 color = "darkorchid2", linewidth = 2, alpha = 0.50) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylim(0,1)+
  ylab("Probability") +
  theme_bw()

ggdat <- tibble(x = 0:3) |>
  mutate(PMF = daltbinom(x,3,0.70,1.5))

PMF.plot.3 <- ggplot() +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_linerange(data = ggdat |> filter(x >= 2),
                 aes(x = x, ymin = 0, ymax = PMF),
                 color = "darkorchid2", linewidth = 2, alpha = 0.50) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylim(0,1)+
  ylab("Probability") +
  theme_bw()

PMF.plot.1 | PMF.plot.2 | PMF.plot.3
```

# Altham's Multiplicative Beta-Binomial Distribution
You may have wondered -- "Can we do the Beta thing with Altham's version of the Binomial Distribution?" The answer is yes!

## Probability Mass Function
The mass function for this distribution is below. 
$$f_X(x) = \left[\int_{0}^1 dp
\left(\frac{1}{c} {n \choose x} p^x (1-p)^{n-x} \theta^{x(n-x)}\right)\left(\frac{\Gamma(\alpha+\beta)}{\Gamma(\alpha)\Gamma(\beta)} p^{\alpha-1}(1-p)^{\beta-1}\right)\right]I(x \in \{0, 1, ..., n\})$$

**Write a function in `R` called `daltbbinom()` for computing the PMF of this distribution. Use your function to plot the distribution where $\alpha=7$ and $\beta=3$ where $\theta=0.50$, $\theta=1$, and $\theta=1.50$. Comment on the differences in the context of a best two-out-of-three match.**\

**Hint**: You can use the `integrate()` in `R` to compute the integral. Note that this requires the function to be vectorized -- it should be unless you have added any block conditionals. 

**Solution**

```{r}
daltbbinom = function(x, n, alpha, beta, theta) {
  # have to vectorize to use integrate
  integrand = Vectorize(function(p){
    i=0:n
    c = sum(choose(n,i)*p^i*(1-p)^(n-i)*theta^(i*(n-i)))
    (1/c)*choose(n,x)*p^x*(1-p)^(n-x)*theta^(x*(n-x))*
      (gamma(alpha+beta)/(gamma(alpha)*gamma(beta)))* 
      p^(alpha-1)*(1-p)^(beta-1)
  })
  integrate(integrand,0,1)$value
}
```

```{r}
#| label: fig-altbbnom-pdf
#| fig-cap: "PMF of Altham's Multiplicative Beta-Binomial Distribution when $\\theta$=0.5, $\\theta$=1 and $\\theta$=1.5"
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

ggdat <- tibble(x=0:3)|>
  rowwise()|>
  mutate(PMF = daltbbinom(x,3,7,3,0.5))

PMF.plot.1 <- ggplot() +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylim(0, 1) +
  ylab("Probability") +
  theme_bw()

ggdat <- tibble(x=0:3)|>
  rowwise()|>
  mutate(PMF = daltbbinom(x,3,7,3,1))

PMF.plot.2 <- ggplot() +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylim(0, 1) +
  ylab("Probability") +
  theme_bw()

ggdat <- tibble(x=0:3)|>
  rowwise()|>
  mutate(PMF = daltbbinom(x,3,7,3,1.5))

PMF.plot.3 <- ggplot() +
  geom_linerange(data = ggdat,
                 aes(x = x, ymin = 0, ymax = PMF)) +
  geom_hline(yintercept = 0) +
  scale_x_continuous("x", breaks = 0:3) +
  ylim(0, 1) +
  ylab("Probability") +
  theme_bw()

PMF.plot.1 | PMF.plot.2 | PMF.plot.3
```

When $\theta=1$, it's a binomial distribution, assuming each leg is independent. When $\theta$ is smaller than 1, the probability of Player A winning the next leg after winning 1 leg is higher, so the probability of winning (WW"L" and WW"W") is significantly higher. When $\theta$ is larger than 1, Player A is more likely to win 2 trials out of 3 (2-1 wins are more likely).

# A Comparison
Compute the probabilities for all outcomes for Player A.

* Loses 0-2 (LL"L", LL"W")
* Loses 1-2 (WLL, LWL)
* Wins 2-0 (WW"W", WW"L")
* Wins 2-1 (LWW, WLW)

**Note**: You'll need to define these outcomes in terms of the random variable. 

**Create a bar plot (or column plot) using `ggplot` to show the probability of each outcome under the following conditions:**

* **Binomial**
* **Negative Binomial**\
**Note**: You can compute the lose probabilities the same way as the win probabilities using success probability ($1-p$)
* **Beta-Binomial**
* **Althman Binomial with $\theta=0.50$ and $\theta=1.5$**
* **Althman Beta-Binomial with $\theta=0.50$ and $\theta=1.5$**

**Comment on the differences in the context of a best two-out-of-three match.**

**Solution**
Defining outcomes in terms of random variable
<!-- need a blank line to recognize bulletpoint -->

* Binomial: 
  * Loses 0-2: $P(X=0) + \frac{1}{3}P(X=1) = f_X(0) + \frac{1}{3}f_X(1)$ 
  * Loses 1-2: $\frac{2}{3}P(X=1)=\frac{2}{3}f_X(1)$
  * Wins 2-0: $P(X=3)+\frac{1}{3}P(X=2) = f_X(3) + \frac{1}{3}f_X(2)$ 
  * Wins 2-1: $\frac{2}{3}P(X=2)=\frac{2}{3}f_X(2)$
* Negative Binomial:
  * Wins 2-0: $P(X=0)=f_X(0)$
  * Wins 2-1: $P(X=1)=f_X(1)$
  * Loses: Using $p=0.30$ instead of $p=0.70$ (the probability of losing the leg), Loses 0-2 is $P(X=0)$ (0 successes before 2 failures) and Loses 1-2 is $P(X=1)$
* Beta-Binomial
  * Loses 0-2: $P(X=0) + \frac{1}{3}P(X=1) = f_X(0) + \frac{1}{3}f_X(1)$
  * Loses 1-2: $\frac{2}{3}P(X=1)=\frac{2}{3}f_X(1)$
  * Wins 2-0: $P(X=3) + \frac{1}{3}P(X=2) = f_X(3) + \frac{1}{3}f_X(2)$
  * Wins 2-1: $\frac{2}{3}P(X=2)=\frac{2}{3}f_X(2)$
* Althman Binomial:
  * Loses 0-2: $P(X=0) + \frac{1}{3}P(X=1) = f_X(0) + \frac{1}{3}f_X(1)$
  * Loses 1-2: $\frac{2}{3}P(X=1)=\frac{2}{3}f_X(1)$
  * Wins 2-0: $P(X=3) + \frac{1}{3}P(X=2) = f_X(3) + \frac{1}{3}f_X(2)$
  * Wins 2-1: $\frac{2}{3}P(X=2)=\frac{2}{3}f_X(2)$
* Althman Beta-Binomial
  * Loses 0-2: $P(X=0) + \frac{1}{3}P(X=1) = f_X(0) + \frac{1}{3}f_X(1)$
  * Loses 1-2: $\frac{2}{3}P(X=1)=\frac{2}{3}f_X(1)$
  * Wins 2-0: $P(X=3) + \frac{1}{3}P(X=2) = f_X(3) + \frac{1}{3}f_X(2)$
  * Wins 2-1: $\frac{2}{3}P(X=2)=\frac{2}{3}f_X(2)$

```{r}
comparing <- tibble(
  outcome=rep(c("Loses 0-2", "Loses 1-2", "Wins 2-0", "Wins 2-1"),times=7),
  distribution=rep(c("Binomial", 
                    "NegBinomial",
                    "BetaBinomial",
                    "Altham(0.5)", 
                    "Altham(1.5)",
                    "AlthamBeta(0.5)",
                    "AlthamBeta(1.5)"),each=4),
  probability=c(
    # Binomial
    dbinom(0,3,0.70)+(1/3)*dbinom(1,3,0.70),
    (2/3)*dbinom(1,3,0.70),
    dbinom(3,3,0.70)+(1/3)*dbinom(2,3,0.70),
    (2/3)*dbinom(2,3,0.70),
    # Negative Binomial 
    dnbinom(0,2,0.30), dnbinom(1,2,0.30), dnbinom(0,2,0.70), dnbinom(1,2,0.70),
    # Beta-Binomial
    dbbinom(0,3,7,3)+(1/3)*dbbinom(1,3,7,3),
    (2/3)*dbbinom(1,3,7,3),
    dbbinom(3,3,7,3)+(1/3)*dbbinom(2,3,7,3),
    (2/3)*dbbinom(2,3,7,3),
    # Altham Binomial theta=0.5
    daltbinom(0,3,0.70,0.5)+(1/3)*daltbinom(1,3,0.70,0.5),
    (2/3)*daltbinom(1,3,0.70,0.5),
    daltbinom(3,3,0.70,0.5)+(1/3)*daltbinom(2,3,0.70,0.5),
    (2/3)*daltbinom(2,3,0.70,0.5),
    # Altham Binomial theta=1.5
    daltbinom(0,3,0.70,1.5)+(1/3)*daltbinom(1,3,0.70,1.5),
    (2/3)*daltbinom(1,3,0.70,1.5),
    daltbinom(3,3,0.70,1.5)+(1/3)*daltbinom(2,3,0.70,1.5),
    (2/3)*daltbinom(2,3,0.70,1.5),
    # Altham Beta-Binomial theta=0.5
    daltbbinom(0,3,7,3,0.5)+(1/3)*daltbbinom(1,3,7,3,0.5),
    (2/3)*daltbbinom(1,3,7,3,0.5),
    daltbbinom(3,3,7,3,0.5)+(1/3)*daltbbinom(2,3,7,3,0.5),
    (2/3)*daltbbinom(2,3,7,3,0.5),
    # Altham Beta-Binomial theta=1.5
    daltbbinom(0,3,7,3,1.5)+(1/3)*daltbbinom(1,3,7,3,1.5),
    (2/3)*daltbbinom(1,3,7,3,1.5),
    daltbbinom(3,3,7,3,1.5)+(1/3)*daltbbinom(2,3,7,3,1.5),
    (2/3)*daltbbinom(2,3,7,3,1.5)
  )
)
```

```{r}
#| label: fig-all-distributions
#| fig-cap: "Probability of each outcome under different distributions"
#| fig-width: 5
#| fig-height: 3
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize
ggplot(comparing) +
  geom_bar(aes (x = distribution,
                y = probability),
            stat="identity") +
geom_hline(yintercept = 0)+
theme_bw()+
labs(x = "Distribution Name",
      y = "Probability") +
facet_wrap(~outcome) +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The probability of losing for Player A is generally lower across distributions, since the probability of winning each leg is 70%. 

* Loses 0-2: This is the case with the lowest probability across distributions, since Player A has 70% chance of winning each leg. It is highest under AlthamBeta ($\theta=0.50$) and BetaBinomial.
* Wins 2-0: Altham($\theta=0.50$) and AlthamBeta($\theta=0.50$) show the highest probability. Clustering effect makes it more likely that the player wins both first 2 legs.
* Loses 1-2 and Wins 2-1: The probability for both outcomes are higher in Altham and AlthamBeta with $\theta=1.50$. It seems like higher theta estimates the alternating outcomes (WLL, LWL, LWW, WLW) to be more likely.
* The probability of winning and losing in the Binomial and NegBinomial are the same ($0.784$).