---
title: Homework 3 -- Regression # Title of the document

# Bibliography Details
bibliography: HW3References.bib  # Name of bib file in the same folder 
bibliographystyle: apa            # Bibliography style

execute:               # Default settings for code blocks
  warning: false       # Do not print warnings
  message: false       # Do not print messages
  size: "scriptsize"   # Set code size to script size

# Below will do html by default. To render the entire project
# Use the following code:
# quarto::quarto_render("HW3.qmd")
format:
  pdf:        # Specify .pdf formatting options 
    documentclass: article     # Document Class
    number-sections: true      # Number Sections
    reference-section-title: "References" # Title for reference section
    code-line-numbers: true    # Line numbers printed in code
    cap-location: top          # Captions at the top of tables 
    geometry:
      - top=0.75in
      - left=0.75in
      - bottom=0.75in
      - right=0.75in
  html:       # Specify HTML formatting options 
    number-sections: true      # Number Sections
    code-fold: true            # Button to show code
    code-line-numbers: true    # Line numbers printed in code
    cap-location: top          # Captions at the top of tables and figures
format-links: [pdf]
---
```{r}
#| echo: false

default_chunk_hook  <- knitr::knit_hooks$get("chunk")

latex_font_size <- c("Huge", "huge", "LARGE", "Large", 
                     "large", "normalsize", "small", 
                     "footnotesize", "scriptsize", "tiny")

knitr::knit_hooks$set(chunk = function(x, options) {
  x <- default_chunk_hook(x, options)
  if(options$size %in% latex_font_size) {
    paste0("\n \\", options$size, "\n\n", 
      x, 
      "\n\n \\normalsize"
    )
  } else {
    x
  }
})

knitr::opts_chunk$set(size = "scriptsize")
```

```{r}
#| label: load-packages
#| include: false
library(tidyverse)
library(patchwork)
```

# Question 1

Below, I simulate a rather quite annoying set of regression data.
```{r}
#| size: "scriptsize"
set.seed(0510)
n <- 250
dat.q1 <- tibble(x1 = rnorm(n, mean = 50, sd = 10),
                 x2 = factor(sample(c("A", "B"), n, replace = TRUE)),
                 x3 = factor(sample(c("Control", "Treat"), n, replace = TRUE)),
                 x4 = factor(sample(c("Low", "High"), n, replace = TRUE), levels=c("Low", "High")),
                 x5 = rnorm(n, mean = 0, sd = 1),
                 x6 = factor(sample(c("Group1", "Group2"), n, replace = TRUE)),
                 x7 = rnorm(n, mean = 10, sd = 2),
                 x8 = rnorm(n, mean = 10, sd = 2)) |>
        mutate(y = 5 +                          # Intercept
                   # Quantiative x1
                   2 * x1 +
                   # Categorical x2
                   (x2 == "B") * 10 +
                   # Categorical x Categorical (x3*x4)
                   (x3 == "Treat") * 5 + (x4 == "High") * 2 + 
                   (x3 == "Treat" & x4 == "High") * 15 +
                   # Quantitative x Categorical  (x5*x6)
                   3 * x5 + (x6 == "Group2") * 4 + 
                   (x5 * (x6 == "Group2")) * 8 + 
                   # Quantitative x Quantitative (x7*x8)
                   1.5 * x7 + 0.5 * x8 + 
                   0.8 * (x7 * x8) + 
                   # Error
                   rnorm(n, mean = 0, sd = 15))
```

## Part a

Fit the regression model.

\begin{align*}
  Y_i =& \beta_0  +
       \beta_1 x_{1i} + \beta_2 I(x_{2i}=\text{B})+ \beta_3 I(x_{3i}=\text{Treat}) + \beta_4 I(x_{4i}=\text{High}) +\\
       & \beta_5 x_{5i} + \beta_6 I( x_{6i}=\text{Group2}) + \beta_7 x_{7i} + \beta_8 x_{8i}+\\
       & \beta_9 (I(x_{3i}=\text{Treat}) ~\times~ I(x_{4i}=\text{High})) + \tag{Categorical $\times$ Categorical}\\
        & \beta_{10} (x_{5i} ~\times~ I(x_{6i}=\text{Group2})) +\tag{Continuous $\times$ Categorical}\\
        & \beta_{11} (x_{7i} ~\times~ x_{8i}) +
         \epsilon_i \tag{Continuous $\times$ Continuous}
\end{align*}

**Solution:** 
First, we fit the regression model using `lm()`.
```{r}
#| size: "scriptsize"
mod.q1 <- lm(y ~ x1 + x2 + x3*x4 + x5*x6 + x7*x8, data=dat.q1)
```

## Part b
Conduct a residual and outlier analysis.

**Solution:**

Looking at @fig-residuals, the histogram of residuals is roughly bell-shaped and the points fit on the line in the qq plot. Therefore, the residuals are normal. The histogram centers at 0, meaning the residuals have a mean of 0. From the scatterplot, we see the residuals spreading equally around 0, suggesting common variance.

In @fig-outlier, we can see that the dataset has several observations that have high leverage. All $D_i$ are below 0.50, so we know that these observations are not strongly influential. 

The scatterplot for the studentized residual and the DFFITs reveal some outliers, however. The residuals for 9 observations fall more than 2 SDs above or below 0. Regarding the DFFITs plot, our model has 10 predictors, so we compare the absolute value of DFFITs to $2*\sqrt{10/250}=0.4$ and $3*\sqrt{10/250}=0.6$. An observation between the 100th and the 125th observation is strongly influential, consistent with the results from studentized residuals calclation. This observation doesn't seem to have high leverage, however.

Overall, the data satisfies the conditions for use of regression regarding residual and outliers.

```{r}
#| eval: false

source("MATH 354/HW3_files/plotResiduals.R")
source("MATH 354/HW3_files/plotInfluence.R")
```

```{r}
source("plotResiduals.R")
source("plotInfluence.R")
```

```{r}
#| label: fig-residuals
#| fig-cap: "Residual Analysis"
#| fig-width: 10
#| fig-height: 5
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

plotResiduals(mod.q1)
```
```{r}
#| label: fig-outlier
#| fig-cap: "Outlider Analysis"
#| fig-width: 10
#| fig-height: 8
#| scale: "80%"
#| fig-align: "center"
#| size: scriptsize

plotInfluence(mod.q1)
```

## Part c
Interpret the intercept, the coefficients on $x_1$ and $x_2$.

**Solution:**
Given the predictors are standardized,

* $\beta_0$ is the expected value of Y when every predictor is at their average (reference value).

* $\beta_1$ is the expected increase in the value of Y with every 1 unit increase in $x_1$, given all other predictors are fixed.

* $\beta_1$ is the expected increase in the value of Y when the reference category for $x_2$ changes from A to B, given all other predictors are fixed.

## Part d
Untangle and interpret the $x_3*x_4$ interaction using algebra. Use @tbl-q1d-emmeans and @fig-q1d-emmeans for context.

**Solution:**
We hold all parts that do not involve $x_3$ and $x_4$ in the model constant, labeled as $C$. Our regression model becomes:
$$Y_i = \beta_3 I(x_{3i}=\text{Treat}) + \beta_4 I(x_{4i}=\text{High}) + \beta_9 (I(x_{3i}=\text{Treat}) ~\times~ I(x_{4i}=\text{High})) + C $$ 
Our expected Y, conditioned on $x_3$ and $x_4$, are as follow:
$$ E(Y_i | x_{3i}=\text{Control}, x_{4i}=\text{Low}) = C $$
$$ E(Y_i | x_{3i}=\text{Control}, x_{4i}=\text{High}) = \beta_4 + C $$
$$ E(Y_i | x_{3i}=\text{Treat}, x_{4i}=\text{Low}) = \beta_3 + C $$
$$ E(Y_i | x_{3i}=\text{Treat}, x_{4i}=\text{High}) = \beta_3 + \beta_4 + \beta_9 + C $$

* $C$ is the expected value of $Y$ in the Control group when condition is Low.

* $\beta_3$ is the expected difference $Y$ between the Control and Treatment group when the condition is Low. In other words, in the Low condition, $\beta_3$ is the expected increase in $Y$ when the reference category changes from Control to Treatment.

* $\beta_4$ is the expected difference $Y$ between the Low and High condition in the Control group.

* In the Treatment condition, when condition changes from Low to High, the expected value of $Y$ increases by $\beta_4 + \beta_9$, meaning it increases by both the expected increase when going from Low to High in the Control group, plus the interaction effect only obtained when high condition and treatment goes together.

```{r}
#| size: "scriptsize"
library(emmeans)
emm.q1d <- emmeans(mod.q1, ~x3*x4) |> # Don't average across levels: use |x2*x6
            data.frame() |>
            rename("Mean"  = "emmean",
                   "Lower" = "lower.CL",
                   "Upper" = "upper.CL")
```

```{r}
#| label: tbl-q1d-emmeans
#| tbl-cap: "Estimated marginal means for the model in Question 1."
#| tbl-align: "center"
#| echo: false
emm.q1d |> 
  knitr::kable(digits = 4)
```

```{r}
#| label: fig-q1d-emmeans
#| fig-cap: "Estimated marginal means for the model in Question 1."
#| fig-width: 5
#| fig-height: 3
#| fig-align: "center"
#| size: "scriptsize"
emm.q1d <- emm.q1d |>
  mutate(Observation = paste("(x3=", x3, ", ", "x4=", x4, ")", sep=""))
ggplot(emm.q1d) + 
  geom_point(aes(x=Mean, y=Observation))+
  geom_errorbar(aes(y=Observation,
                    xmin=Lower, xmax=Upper),
                width = 0.25)+
  # facet_wrap(~x2*x6)+
  theme_bw()+
  xlab("Estimated Mean of Y")
```

## Part e
Untangle and interpret the $x_5*x_6$ interaction using algebra. Use @tbl-q1e-emtrends and @fig-q1e-emtrends for context.

**Solution:**
We hold all parts that do not involve $x_5$ and $x_6$ in the model constant, labeled as $C$. Our regression model becomes:
$$Y_i = \beta_5 x_{5i} + \beta_6 I( x_{6i}=\text{Group2}) + \beta_{10} (x_{5i} ~\times~ I(x_{6i}=\text{Group2})) + C $$ 
Our expected Y, conditioned on $x_5$ and $x_6$, are as follow:

* Group 1: 
$$Y_i = \beta_5 x_{5i} + C $$
    * $C$ is the expected value of $Y$ when $x_5=0$.
    
    * $\beta_5$ is the increase in the expected value of $Y$ when $x_5$ increases by 1 unit.

* Group 2: 
$$Y_i = (\beta_5 + \beta_{10})x_{5i} + (\beta_6 + C) $$

    * $\beta_6 + C$ is the expected value of $Y$ when $x_5=0$.

    * $\beta_5 + \beta_10$ is the increase in the expected value of $Y$ when $x_5$ increases by 1 unit.

```{r}
#| size: "scriptsize"
emm.q1e <- emtrends(mod.q1, var="x5", ~x5|x6) |> 
            data.frame() |>
            rename("Slope"  = "x5.trend",
                   "Lower" = "lower.CL",
                   "Upper" = "upper.CL") |>
            dplyr::select(x6, Slope, SE, df, Lower, Upper)
```
```{r}
#| label: tbl-q1e-emtrends
#| tbl-cap: "Estimated marginal trends for the model in Question 1."
#| tbl-align: "center"
#| echo: false
emm.q1e |> 
  knitr::kable(digits = 4)
```

```{r}
#| label: fig-q1e-emtrends
#| fig-cap: "Estimated marginal trends for the model in Question 1."
#| fig-width: 5
#| fig-height: 3
#| fig-align: "center"
#| size: "scriptsize"
emm.q1e <- emmeans(mod.q1, ~x5*x6,
                    at = list(x5 = seq(-3, 3, length.out=500))) |> # Don't average across levels: use |x2*x3*x4
  data.frame() |>
            rename("Mean"  = "emmean",
                   "Lower" = "lower.CL",
                   "Upper" = "upper.CL")

ggplot(emm.q1e) + 
  geom_line(aes(x=x5, y=Mean, color=x6))+
  geom_ribbon(aes(x=x5, ymin=Lower, ymax=Upper, fill=x6), alpha=0.25)+
  # facet_wrap(~x2*x3*x4)+
  theme_bw()
```


## Part f
Untangle and interpret the $x_7*x_8$ interaction using algebra. Use @tbl-q1f-emtrends-x7 and @tbl-q1f-emtrends-x8 for context.

**Solution:**
We hold all parts that do not involve $x_7$ and $x_8$ in the model constant, labeled as $C$. Our regression model becomes:
$$Y_i = \beta_7 x_{7i} + \beta_8 x_{8i} + \beta_{11} (x_{7i} ~\times~ x_{8i}) + C $$ 
We can combine like terms to get:

$$Y_i = (\beta_7 + \beta_{11}x_{8i})x_{7i} + \beta_8 x_{8i} + C $$ 
$$Y_i = (\beta_8 + \beta_{11}x_{7i})x_{8i} + \beta_7 x_{7i} + C $$ 

* $\beta_7 + \beta_{11}x_{8i}$ is the change in $Y$ for a unit change in $x_7$ after adjusting for the
simultaneous linear change in $x_8$.

* $\beta_8 + \beta_{11}x_{7i}$ is the change in $Y$ for a unit change in $x_8$ after adjusting for the
simultaneous linear change in $x_7$.

* $\beta_11$ is the change in the **slope of $x_7$** when $x_8$ increases by 1 unit, and is also the change in the slope of $x_8$ when $x_7$ increases by 1 unit. From the tables, we can see that increases in $x_8$ is associated with increases in the slope of $x_7$, and vice versa.

* When $x_8=0$, $\beta_7$ is the change in the expected value of $Y$ when $x_7$ increases by 1 unit.

* When $x_7=0$, $\beta_8$ is the change in the expected value of $Y$ when $x_8$ increases by 1 unit.

<!-- When we take partial derivatives (which tells us how a multivariable function changes when one variable changes and others are fixed), we get the following functions:

$$\frac{\partial Y_i}{\partial x_{7}} = \beta_7 + \beta_{11}x_{8i}$$
$$\frac{\partial Y_i}{\partial x_{8}} = \beta_{8} + \beta_{11}x_{7i}$$
The expected value of $Y$ increases by $\beta_{7} + \beta_{11}x_{8i}$ when $x_7$ increases by 1 unit, after adjusting for the simultaneous linear change in $x_8$. The expected value of $Y$ increases by $\beta_8 + \beta_{11}x_{8i}$ when $x_8$ increases by 1 unit, after adjusting for the simultaneous linear change in $x_7$. -->

```{r}
#| size: "scriptsize"
emm.q1f.x7 <- emtrends(mod.q1, var="x7", ~x7|x8, 
                    at = list(x8 = mean(dat.q1$x8) + c(-1, 0 ,1)*sd(dat.q1$x8))) |> 
            data.frame() |>
            rename("Slope"  = "x7.trend",
                   "Lower" = "lower.CL",
                   "Upper" = "upper.CL") |>
            dplyr::select(x8, Slope, SE, df, Lower, Upper)
```

```{r}
#| label: tbl-q1f-emtrends-x7
#| tbl-cap: "Estimated marginal trends for the model in Question 1."
#| tbl-align: "center"
#| echo: false
emm.q1f.x7 |> 
  knitr::kable(digits = 4)
```

```{r}
#| size: "scriptsize"
emm.q1f.x8 <- emtrends(mod.q1, var="x8", ~x8|x7, 
                    at = list(x7 = mean(dat.q1$x7) + c(-1, 0 ,1)*sd(dat.q1$x7))) |> 
            data.frame() |>
            rename("Slope"  = "x8.trend",
                   "Lower" = "lower.CL",
                   "Upper" = "upper.CL") |>
            dplyr::select(x7, Slope, SE, df, Lower, Upper)
```
```{r}
#| label: tbl-q1f-emtrends-x8
#| tbl-cap: "Estimated marginal trends for the model in Question 1."
#| tbl-align: "center"
#| echo: false
emm.q1f.x8 |> 
  knitr::kable(digits = 4)
```

A Johnson Neyman analysis using the `interactions` package for `R` [@interactions], shows that the slope of $x_7$ is statistically discernible for all observed values of $x_8$ (@fig-q1f-jni-x7), and vice versa (@fig-q1f-jni-x8).

```{r}
#| label: fig-q1f-jni-x8
#| fig-cap: "Johnson Neyman interval for the model in Question 1."
#| fig-width: 5
#| fig-height: 3.5
#| fig-align: "center"
#| size: "scriptsize"
library(interactions)
johnson_neyman(mod.q1, pred = x7, modx = x8, control.fdr = T)
```

```{r}
#| label: fig-q1f-jni-x7
#| fig-cap: "Johnson Neyman interval for the model in Question 1."
#| fig-width: 5
#| fig-height: 3.5
#| fig-align: "center"
#| size: "scriptsize"
johnson_neyman(mod.q1, pred = x8, modx = x7, control.fdr = T)
```

## Part g
Standardize the continuous data and refit the regression model. Redo part f. What do you notice about the coefficient estimates and why does this make the untangling process easier? 

**Solution:** 
First, we fit the regression model using `lm()`.
```{r}
#| size: "scriptsize"
# Standardize Data
dat.z <- dat.q1 |>
  mutate(across(c(x1, x5, x7, x8), ~ as.vector(scale(.))))
# Refit Model
mod.q1z <- lm(y ~ x1 + x2 + x3*x4 + x5*x6 + x7*x8, data=dat.z)
```

```{r}
#| size: "scriptsize"
emm.q1g.z7 <- emtrends(mod.q1z, var="x7", ~x7|x8,
                    at = list(x8 = c(-1, 0 ,1))) |>
          data.frame() |>
          rename("Slope"  = "x7.trend",
                 "Lower" = "lower.CL",
                 "Upper" = "upper.CL") |>
          dplyr::select(x8, Slope, SE, df, Lower, Upper)
```
```{r}
#| label: tbl-q1-emtrends-z7
#| tbl-cap: "Estimated marginal trends for the model in Question 1 when quantitative variables are standardized."
#| tbl-align: "center"
#| echo: false
emm.q1g.z7 |> 
          knitr::kable(digits = 2, row.names = F)
```
```{r}
#| size: "scriptsize"
emm.q1g.z8 <- emtrends(mod.q1z, var="x8", ~x8|x7,
                      at = list(x7 = c(-1, 0 ,1)))|>
              data.frame() |>
              rename("Slope"  = "x8.trend",
                    "Lower" = "lower.CL",
                    "Upper" = "upper.CL") |>
              dplyr::select(x7, Slope, SE, df, Lower, Upper)

```
```{r}
#| label: tbl-q1-emtrends-z8
#| tbl-cap: "Estimated marginal trends for the model in Question 1 when quantitative variables are standardized."
#| tbl-align: "center"
#| echo: false
emm.q1g.z8 |> 
          knitr::kable(digits = 2, row.names = F)
```

After transformation,transformation

* $\beta_7$ is the effect of a 1-SD change in $x_7$ when $x_8$ is at its mean.

* $\beta_7$ is the effect of a 1-SD change in $x_7$ when $x_8$ is at its mean.

* $\beta_11$ now becomes the change in the slope of 1 standardized predictor when the other predictor shifts by 1 standard deviation. For example, looking at the effect of $x_7$:

$$Y_i = (\beta_7 + \beta_{11}x_{8i})x_{7i} + \beta_8 x_{8i} + C $$ 

* When $z_8=-1$, $$Y_i = (\beta_7 - \beta_{11})z_{7i} - \beta_8 + C $$. In other words, when $x_8$ is 1 SD below the means, for every increase in the SD of $x_7$, the expected value of $Y$ increases by $\beta_7 - \beta_{11}$. When $x_7$ is at average, the expected value of $Y$ is $C - \beta_8$.
This transformation makes the untangling process easier. In @tbl-q1-emtrends-z7, in the slope column, the second row is the estimated value for $\beta_7 - \beta_{11}$, the third row is the estimated value for $\beta_7 + \beta_{11}$, and the difference between rows is the estimated value for $\beta_{11}$.

# Question 2
@Cui25 investigate how freezing and cryogenic temperatures affect the strength, flexibility, stiffness, and internal breaking behavior of 3D-printed CF/PEEK composites.

Carbon fiber reinforced PEEK (CF/PEEK) is a high-performance composite used in 3D printing for aerospace, deep-sea, polar research. The researchers aimed to assess whether and how extremely cold environments impact this composite to understand how 3D-printed CF/PEEK performs in freezing conditions. This work would be important to understand its limits and develop preventative methods to protect against structural failures.

The researchers created fifteen 3D printed dog-bone-shaped test specimens and placed them in a chamber cooled with liquid nitrogen to $25^\circ\text{C}$, $-25^\circ\text{C}$, $-75^\circ\text{C}$, $-125^\circ\text{C}$, and $-175^\circ\text{C}$. To test strength of the specimen, a machine pulled both ends of the specimen until it snapped. 

In the table below, the researchers record the average tensile strength for three specimens from each temperature level.

| Temperature          | Mean Tensile Strength | 
| :--                  | :--                   |
| $25^\circ\text{C}$   | 81.93 MPa             |
| $-25^\circ\text{C}$  | 76.63 MPa             |
| $-75^\circ\text{C}$  | 63.23 MPa             |
| $-125^\circ\text{C}$ | 32.13 MPa             |
| $-175^\circ\text{C}$ | 12.83 MPa             |

## Part a
Compute the estimates we derived using calculus in lecture 16: 
\begin{align*} 
\widehat{\beta}_0 &= \bar{y} - \widehat{\beta}_1 \bar{x}_1 \\ 
\widehat{\beta}_1 &= \frac{\sum_{i=1}^n (x_{1i} - \bar{x}_1)(y_i-\bar{y})}{\sum_{i=1}^n (x_{1i}-\bar{x}_1)^2}.
\end{align*}

## Part b
Compute the estimates we derived using linear algebra in lecture 16: 
$$\widehat{\boldsymbol{\beta}} = \left(\mathbf{X}'\mathbf{X}\right)^{-1} \mathbf{X}'\mathbf{Y}.$$
**Hint:** Run `?matrix()` in `R` for a quick reminder about how to create a matrix. 



## Part c
Compute the maximum likelihood estimates using `optim()` in `R` as described in lecture 16:
$$\max_{\beta_0, \beta_1, \sigma} \mathcal{L}(\boldsymbol{\theta}=\{\beta_0, \beta_1, \sigma\}).$$

## Part d
Fit the model using `lm()` and plot the resulting model. Compare to the results from parts (a) through (c). 

# Question 3
@Bracht16 evaluate MFAP4 as a biomarker for severity of fibrosis in hepatitic C patients. 

```{r}
#| size: "scriptsize"
dat.mfap4 <- read_csv("Bracht2016.csv")
```

## Part a
Fit the model predicting `MFAP4` using `Fibrosis.Stage`. Write out the resulting estimate model and interpret the coefficients.

## Part b
Conduct a residual analysis on the model in part a to demonstrate how to identify when you need a log transformation.


## Part c
Fit the model predicting `log(MFAP4)` using `Fibrosis.Stage`. Use the natural log, and not the $\log_2(\cdot)$ recorded in the data file. Write out the resulting estimate model and interpret the coefficients.


## Part d
Conduct a residual analysis on the model in part c to demonstrate that the conditions for using regression are met.

## Part e
Fit the model predicting `log(MFAP4)` using `Fibrosis.Stage`, `Age`, and `Gender`. Use the natural log, and not the $\log_2(\cdot)$ recorded in the data file. Write out the resulting estimate model and interpret the coefficients.



# Question 4

## Part a
Simulate data to match regression assumptions in a loop that runs $R=1000$ times.
\begin{align*}
    x_i &\sim \text{Uniform}(a=0,b=10)\\
    \epsilon_i &\sim \text{Gaussian}(\mu=0, \sigma=1)\\
    y_i &= 2 + 1.5 x_i + \epsilon_i\\
    n &= 30
\end{align*}
In each iteration, fit a regression line to the data and store the coefficient estimates. Plot the resulting estimates. What do you notice?

## Part b
Simulate data to match regression assumptions in a loop that runs $R=1000$ times.
\begin{align*}
    x_i &\sim \text{Uniform}(a=0,b=10)\\
    \epsilon_i &\sim \text{Shifted Exponential}(\lambda=1, \mu=0)\\
    y_i &= 2 + 1.5 x_i + \epsilon_i\\
    n &= 30
\end{align*}
In each iteration, fit a regression line to the data and store the coefficient estimates. Plot the resulting estimates. What do you notice?

**Hint:** Use the `tolerance::r2exp()` function with `rate=1` and `shift=-1`.

## Part c
Simulate data to match regression assumptions in a loop that runs $R=1000$ times.
\begin{align*}
    x_i &\sim \text{Uniform}(a=0,b=10)\\
    \epsilon_i &\sim \text{Cauchy}(m=0, s=1)\\
    y_i &= 2 + 1.5 x_i + \epsilon_i\\
    n &= 30
\end{align*}
In each iteration, fit a regression line to the data and store the coefficient estimates. Plot the resulting estimates. What do you notice?

